
\documentclass{article}

\usepackage{url}
\usepackage{parskip} 	

%other packages for formatting
\RequirePackage{color}
\RequirePackage{graphicx}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[margin=.8in]{geometry}

%tabularx environment
\usepackage{tabularx}

%for lists within experience section
\usepackage{enumitem}

% centered version of 'X' col. type
\newcolumntype{C}{>{\centering\arraybackslash}X} 

%to prevent spillover of tabular into next pages
\usepackage{supertabular}
\usepackage{tabularx}
\newlength{\fullcollw}
\setlength{\fullcollw}{0.47\textwidth}

%custom \section
\usepackage{titlesec}				
\usepackage{multicol}
\usepackage{multirow}

%CV Sections inspired by: 
%http://stefano.italians.nl/archives/26
\titleformat{\section}{\Large\scshape\raggedright}{}{0em}{}[\titlerule]
\titlespacing{\section}{0pt}{10pt}{10pt}

%for publications
\usepackage[style=authoryear,sorting=ynt, maxbibnames=2]{biblatex}

%Setup hyperref package, and colours for links
\usepackage[unicode, draft=false]{hyperref}
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour}
\addbibresource{citations.bib}
\setlength\bibitemsep{1em}

%for social icons
\usepackage{fontawesome5}

% job listing environments
\newenvironment{jobshort}[2]
{
	\begin{tabularx}{\linewidth}{@{}l X r@{}}
		\textbf{#1} & \hfill &  #2 \\[3.75pt]
	\end{tabularx}
}
{
}

\newenvironment{joblong}[2]
{
	\begin{tabularx}{\linewidth}{@{}l X r@{}}
		\textbf{#1} & \hfill &  #2 \\[3.75pt]
	\end{tabularx}
	\begin{minipage}[t]{\linewidth}
		\begin{itemize}[nosep,after=\strut, leftmargin=1em, itemsep=3pt,label=--]
		}
		{
		\end{itemize}
	\end{minipage}    
}



%----------------------------------------------------------------------------------------
%	BEGIN DOCUMENT
%----------------------------------------------------------------------------------------
\begin{document}
	
	% non-numbered pages
	\pagestyle{empty} 
	
	%----------------------------------------------------------------------------------------
	%	TITLE
	%----------------------------------------------------------------------------------------
	
	% \begin{tabularx}{\linewidth}{ @{}X X@{} }
		% \huge{Your Name}\vspace{2pt} & \hfill \emoji{incoming-envelope} email@email.com \\
		% \raisebox{-0.05\height}\faGithub\ username \ | \
		% \raisebox{-0.00\height}\faLinkedin\ username \ | \ \raisebox{-0.05\height}\faGlobe \ mysite.com  & \hfill \emoji{calling} number
		% \end{tabularx}
	
	\begin{tabularx}{\linewidth}{@{} C @{}}
		\Huge{Jeffrey M Epstein} \\[7.5pt]
		\href{https://jeffreymepstein.github.io/}{\raisebox{-0.05\height}\faGlobe \ jeffreymepstein.github.io} \ $|$ \ 
		\href{mailto:jeffrey.m.epstein@gmail.com}{\raisebox{-0.05\height}\faEnvelope \ jeffrey.m.epstein@gmail.com} \ $|$ \  {\raisebox{-0.05\height}\faMapMarker  \ San Francisco, CA, USA}
		
	\end{tabularx}
	
	%----------------------------------------------------------------------------------------
	% EXPERIENCE SECTIONS
	%----------------------------------------------------------------------------------------
	

	\section{Work Experience}
	
	\begin{joblong}{Atom Computing, Senior Applications Engineer}{Aug 2021 - present}
\item Decoder implementation and development of tools for logical circuit compilation and logical error rate analysis, suitable for comparing error correction schemes in preparation for small-scale experimental demonstrations
\item Development of circuit simulation and statistical analysis tools for efficient gate characterization
\item Led characterization/benchmarking component of DARPA US2QC program
\item Development of novel state preparation algorithm for constrained optimization

\item Built tools based on Q-CTRL software for optimization of pulse sequences on atomic platform,
facilitating design of rapid gates robust against various sources of noise.

\item Supervised company’s first theory intern, leading to her authorship on a scientific publication.
\end{joblong}
	
	
\begin{joblong}{National Institute of Standards and Technology, Postdoctoral Scholar}{Feb 2021 - June 2021}
		\item Research on quantum information and thermodynamics, leading to publication.
		\item NIST NRC Postdoctoral Research Associateship, 2021
\end{joblong}
	
	
\begin{joblong}{University of California, Berkeley, Graduate Student Researcher}{June 2015 - Dec 2020}
\item Research and publication on problems in quantum information theory and nonequilibrium statistical mechanics.
\item National Defense Science and Engineering Graduate (NDSEG) Fellowship, 2016 - 2021
\end{joblong}

\begin{joblong}{University of California, Berkeley, Graduate Student Instructor}{Sep 2014 - May 2015}
\item Physics 112 (intro. to statistical and thermal physics), Physics 7b (intro. thermodynamics and electromagnetism for scientists and engineers). Taught sections, held regular
office hours, and graded problem sets and exams.
\end{joblong}

\begin{joblong}{IBM Research, TJ Watson Research, Quantum Computing Intern}{Sep 2012 - June 2013}
\item Studied robustness of randomized benchmarking (RB) under varying noise models, leading to a highly-cited publication used in the field as evidence for the validity of RB for benchmarking quantum processors subject to realistic physical noise.
\end{joblong}
	

	\section{Education}
	\begin{tabularx}{\linewidth}{@{}l X@{}}	
		2014-2020 & University of California, Berkeley, PhD in physics\\
		& Dissertation: Statistical Mechanics of Transport Processes in Active Matter\\
		\\		
		2013-2014 & Perimeter Institute for Theoretical Physics, MSc in Theoretical Physics (PSI)\\
		\\
		2008-2012 & Harvard College, AB magna cum laude with high honors in field\\
		& Concentration in Chemistry and Physics\\
		& Secondary field in Mathematics\\
		& Language citation in Chinese
	\end{tabularx}
	
	

	
	
	
	%----------------------------------------------------------------------------------------
	%	PUBLICATIONS
	%----------------------------------------------------------------------------------------
	\section{Publications}
	\begin{refsection}[citations.bib]
		\nocite{*}
		\printbibliography[heading=none]
		\begin{enumerate}
			
			\item Repeated ancilla reuse for logical computation on a neutral atom quantum computer
			JA Muniz, D Crow, H Kim, JM Kindem, WB Cairncross, A Ryou, ...
			arXiv preprint arXiv:2506.09936
			3	2025
			\item High-Fidelity Universal Gates in the  Ground-State Nuclear-Spin Qubit
			JA Muniz, M Stone, DT Stack, M Jaffe, JM Kindem, L Wadleigh, ...
			PRX Quantum 6 (2), 020334
			30	2025
			\item Reasoning across spacelike surfaces in the Frauchiger-Renner thought experiment
			\textbf{JM Epstein}
			arXiv preprint arXiv:2503.00966
			2025
			
			\item  Thermally driven quantum refrigerator autonomously resets superconducting qubit. M Aamir, P Suria, J Guzmán, C
			Castillo-Moreno, \textbf{JME}, N Yunger Halpern, S Gasparinetti. Nature Physics 21 (2), 318-323
			54	2025
			
			\item Fault-tolerant quantum computation with a neutral atom processor
			BW Reichardt, A Paetznick, D Aasen, I Basov, JM Bello-Rivas, ...
			arXiv preprint arXiv:2411.11822
			6	2024
			\item Logical computation demonstrated with a neutral atom quantum processor
			BW Reichardt, A Paetznick, D Aasen, I Basov, JM Bello-Rivas, ...
			arXiv e-prints, arXiv: 2411.11822
			
			
		\item Note on simple and consistent gateset characterization including calibration and decoherence errors. \textbf{JME}.
		arXiv:2402.17727 (2024)
		\item Subspace Correction for Constraints. K Pawlak, \textbf{JME}, D Crow, S Gandhari, M Li, T Bohdanowicz, J King.
		arXiv:2310.20191 (2024)
		\item Iterative assembly of 171Yb atom arrays in cavity-enhanced optical lattices. M Norcia et al. arXiv:2401.16177 (2024)
		\item  Mid-circuit qubit measurement and rearrangement in a 171Yb atomic array. M Norcia et al. arXiv:2305.19119 (2023)
	
		\item  Odd Diffusivity of Chiral Random Motion. C Hargus, \textbf{JME}, KK Mandadapu. Phys. Rev. Lett. 127, 178001 (2021).
		\item  Quantum noise limits for a class of nonlinear amplifiers. \textbf{JME}, KB Whaley, J Combes. Phys. Rev. A 103 (5), 052415
		(2021).
		
			\item Continuous quantum error correction for evolution under time-dependent Hamiltonians. J Atalaya, S Zhang, MY Niu,
		A Babakhani, HCH Chan, \textbf{JME}, KB Whaley. Phys. Rev. A 103, 042406 2021
		
		\item  Time reversal symmetry breaking and odd viscosity in active fluids: Green-Kubo and NEMD results. C Hargus, K Klymko,
		\textbf{JME}, KK Mandadapu. J. Chem. Phys. 152, 201102 (2020).
		\item  Time reversal symmetry breaking in two-dimensional non-equilibrium viscous fluids. \textbf{JME}, KK Mandadapu.
		Phys. Rev. E 101, 052614 (2020).
		\item  Statistical Mechanics of Transport Processes in Active Fluids II: Equations of Hydrodynamics for Active Brownian Particles. \textbf{JME}, K Klymko, KK Mandadapu. J. Chem. Phys. 150, 164111 (2019).
		\item  Postponing the orthogonality catastrophe: efficient state preparation for electronic structure simulations on quantum
		devices. NM Tubman, C Mejuto-Zaera, \textbf{JME}, D Hait, DS Levine, W Huggins, Z Jiang, JR McClean, R Babbush, M HeadGordon, KB Whaley. arXiv:1809.05523 (2018).
		\item  Quantum Speed Limits for Quantum Information Processing Tasks. \textbf{JME}, KB Whaley. Phys. Rev. A 95, 042314 (2017).
		\item  Investigating the Limits of Randomized Benchmarking Protocols. \textbf{JME}, AW Cross, E Magesan, and JM Gambetta. Phys.
		Rev. A 89, 062321 (2014)
		\item  CD36 in the periphery and brain synergize in stroke injury in hyperlipidemia. E Kim, M Febbraio, Y Bao, AT Tolhurst,
		\textbf{JME}, S Cho. Annals of Neurology. 71(6) (2012)
	\end{enumerate}
	\end{refsection}
	
	%----------------------------------------------------------------------------------------
	%	SKILLS
	%----------------------------------------------------------------------------------------

	
	\vfill
	\center{\footnotesize Last updated: \today}
	
\end{document}
